Topological Equivariant Artist Model

### 2021-03-24 slides/main.tex
figure tweaks
moved figures creation code into a mostly central folder, added k labels to fiber bundle fig
bulk changes, everything but timeline

\documentclass[xcolor={dvipsnames}, handout]{beamer}

%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}

@@

\usepackage{tabulary}
\usepackage{multirow}
\usepackage{graphicx}

@@

\usepackage[backend=bibtex, style=numeric-comp, sorting=none, doi=false,isbn=false,url=false, giveninits=true]{biblatex}
\bibliography{../draft/glasslab_viz.bib}

@@

\begin{frame}{Visualizations are structure preserving maps}
    \begin{figure}
        \includegraphics[width=1\linewidth]{figures/intro/dar.png}
    \end{figure}
    The aim of this work is to rearchitecture Matplotlib
    \pause 
    to take advantage of developments in software design, data structures, and visualization
    \pause
    to improve consistency, reusability, and discoverability,
    \pause
    so domain specific tool developers can build structure preserving visualization tools.
\end{frame}

\begin{frame}{Visualization component constraints}
    \begin{description}
        \item[equivariance] properties of data and visual encoding match
        \item[continuity] connectivity of data and visual encoding match
        \item[composibility] structure preserved by individual components is preserved in combined components
    \end{description}
\end{frame}

\subsection{Background}
\begin{frame}{Tools are tuned to date continuity \cite{HeerSoftware2006}}
    \begin{columns}
            \column{0.33\textwidth}
            \begin{figure}
                \includegraphics[width=1\textwidth]{figures/intro/data_formatting.png}
                \caption{Based on fig~2.5 in Munzner's VAD\cite{munznerVisualizationAnalysisDesign2014}}
            \end{figure}
            \begin{enumerate}
                \item Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}
                \item ggplot\cite{wickhamGgplot2ElegantGraphics2016a}
                \item Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, Altair\cite{vanderplasAltairInteractiveStatistical2018}
            \end{enumerate}
            \pause
            \column{0.33\textwidth}
            \begin{figure}
                \includegraphics[width=1\textwidth]{figures/intro/landsat.png}
            \end{figure}
            \begin{enumerate}
                \item ImageJ\cite{schneiderNIHImageImageJ2012}, ImagePlot\cite{studiesCulturevisImageplot2021}
                \item Napari\cite{nicholas_sofroniew_2021_4533308}
            \end{enumerate}
            \pause
            \column{0.33\textwidth}
            \begin{figure}
                \includegraphics[width=1\textwidth]{figures/intro/dataset_diagram.png}
                \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
            \end{figure}
            \begin{enumerate}
                \item Gephi\cite{bastianGephiOpenSource2009}
                \item Graphviz\cite{ellsonGraphvizOpenSource2002}
                \item Networkx\cite{HagbergExploringNetwork2008}
            \end{enumerate}
    \end{columns}
\end{frame}

\begin{frame}{Visualizations are tuned to data continuinity\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007}, 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveci2012vtk}, MayaVi\cite{RamachandranMayaVI2011}, ParaView\cite{ahrens2005paraview}, Titan\cite{brianwylieUnifiedToolkitInformation2009}
    \end{enumerate}
\end{frame}

\begin{frame}{Structure is encoded in variables and continuity}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Image is figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
    \begin{description}
        \item[binding] metadata are structural \textit{keys} with associated \textit{values} (Munzner \cite{munznerVisualizationAnalysisDesign2014})] 
        \item[continuity] Fiber bundles can be a common data abstraction (Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989})
        \item[variables] Fibers can hold schema like encodings of variables (Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES})
    \end{description}
\end{frame}

\begin{frame}{Visualizations are (mostly) evaluated on equivariance}
    \begin{description}
        \item[Expressiveness] structure preserving mappings from data to graphic (Mackinlay \cite{mackinlayAutomatingDesignGraphical1986})
        \item[Effectiveness] design choices made in deference to perceptual saliency (Mackinlay \cite{clevelandResearchStatisticalGraphics1987,clevelandGraphicalPerceptionTheory1984,chambersGraphicalMethodsData1983a, munznerVisualizationAnalysisDesign2014})
        \item[Naturalness] easier to understand when properties match (Norman \cite{norman_things_smart})
        \item[Graphical Integrity] graphs show \textbf{only} the data (Tufte \cite{tufteVisualDisplayQuantitative2001})
    \end{description}
\end{frame}

\begin{frame}{Models describe composition}
    \begin{description}
        \item[language model] APT, GoG: syntax, semantics, and grammar of graphics (Mackinlay, Wilkenson  \cite{mackinlayAutomatingDesignGraphical1986, mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005})
        \item[functional dependencies] constrained maps between data and visual representation(Sugibuchi \cite{sugibuchiFramwork2009}) 
        \item[category theory] the semiotics of visualization are commutative (Vickers \cite{vickersUnderstandingViz2013})
        \item[algebraic process] data ($\alpha$) and viz ($\omega$) transforms are symmetric (Kindlmann and Scheidegger \cite{kindlmannAlgebraicProcessVisualization2014})
        \begin{columns}
            \column{.5\textwidth}

            \begin{description}
                \item[D] data 
                \item[R] representations
                \item[V] visualizations
            \end{description}
            \column{.5\textwidth}
            \begin{equation*}
                \begin{tikzcd}[ampersand replacement=\&]
                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
                \end{tikzcd}
                \end{equation*}

@@

\begin{frame}{Topological Equivariant Artist Model}
    The Artist $\mathscr{\vartist}$ is a map from data $\mathscr{\dtotal}$ to graphic $\mathscr{\gtotal}$
    \begin{equation}
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation}
    \pause
     that carries a homomorphism of monoid actions 
    \begin{equation}
    \varphi: \monoid \rightarrow \monoid^{\prime}
    \end{equation}
    \pause
    such that artists are equivariant maps 
    \begin{equation}
    \mathscr{\vartist}(m\cdot \delement) = \varphi(m)\cdot\mathscr{\vartist}(\delement) 
    \end{equation}
    \pause 
    with a deformation retraction from graphic to data space. 
    \note{some kinda words about the continuity}
\end{frame}

\begin{frame}{Data Bundle}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/fiberbundle.png}
    \end{figure}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the projection map $\pi$
    \begin{equation}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation}
    where \dtotal\ is the total data space, \dfiber\ is the variable space, and \dbase\ encodes the continuity.
\end{frame}

\begin{frame}{Variables: Fiber}
    Given a space of all possible values \ftotal\
    \begin{equation}
        \label{eq:data_types}
        \begin{tikzcd}[ampersand replacement=\&]
            \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] \& \ftotal \arrow[d, "\pi"] \\
            \fnames \arrow[r, "\fsection"']                  \& \ftypes       
        \end{tikzcd}
    \end{equation}
    a fiber component is the restricted space $\ftotal_{\fsection(\fname)}$. 
    \begin{equation}
        \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
    \end{equation}
    \begin{description}
        \item[\ftypes] data types of the variables in the dataset 
        \item[\ftotal] disjoint union of all values of type $\ftype \in \ftypes$ 
        \item[\fnames] variable names, $\fname \in \fnames$
        \item[\fttype] \ftotal\ restricted to the data type of a named variable   
    \end{description}
\end{frame}

\begin{frame}{Variable types are dimensions of the fiber}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/fiber.png}
        \caption{}
        \label{fig:data_fiber_example}
    \end{figure}
    \begin{description}
        \item[plane]  $\dfiber=\reals\times\reals$, (time, temperature)
        \item[cube]  $\reals\times\realsp\times\reals$, (time, wind=(speed, direction))
    \end{description}
\end{frame}

\begin{frame}{Structure of Components: Monoid \& Monoid Actions}
A monoid \monoid\ is a set with
\begin{description}
    \item[associative binary operator] $\ast:\monoid \times \monoid\rightarrow \monoid$
    \item[identity element] $e\in \monoid$ such that $e\ast a= a \ast e = a$ for all $a \in \monoid$. 
\end{description}
\pause
\begin{block}{left monoid action}
A set \dfiber\ with an action\cite{nlab:action} $\bullet: \monoid\times \dfiber \rightarrow \dfiber$ with the properties:
    \begin{align*}
        \textbf{associativity}\;& \text{for all } f,g \in \monoid \text{ and } x\in \dfiber,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
        \textbf{identity}\;& \text{for all } x\in \dfiber, e\in \monoid,\,  e\bullet x = x 
    \end{align*}
\end{block}
\end{frame}

\begin{frame}{Monoid Actions: Permutation}
    \begin{figure}
        \includegraphics[width=1\linewidth]{figures/math/monoid_emoji.png}
    \end{figure}
\end{frame}

\begin{frame}{Why monoids? partial orders}
    \begin{figure}
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/monoid_hasse.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/monoid_monotone.png}
            \onslide<3>\includegraphics[width=1\linewidth]{figures/math/monoid_maps.png}
        \end{overprint}
    \caption{Inspired by definition 1.59 diagram in Spivak and Fong's An Invitation to Applied Category Theory \cite{fongInvitationAppliedCategory2019}}
    \end{figure}
\end{frame}

\begin{frame}{Data Continuity: Base space}
    \begin{figure}[H]
    \includegraphics[height=.5\textheight]{figures/math/k_qspace.png}
    \label{fig:base_space_div}
\end{figure}
where the total space can be decomposed into components 
\begin{equation}
    \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
\end{equation}
\end{frame}

\begin{frame}{Data connectivity is encoded as the base space}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
        \caption{}
        \label{fig:base_example}
    \end{figure}
    \begin{description}
        \item[points] data is 0D discrete
        \item[line]  data is lies on the 1D continuous interval \dbase\
    \end{description}
\end{frame}

\begin{frame}{Values: Section}
    For any fiber bundle, there exists a map
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$.  Set of all global sections is denoted $\Gamma(\dtotal)$.
     \pause
     \begin{block}{Record}
     Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber.
\end{block}
\end{frame}

\begin{frame}{Sample dataset}
    \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
    \begin{itemize}
        \item \dfiber\ is $\reals\times\reals$
        \item \dbase\ is interval $\left[0,1\right]$
        \item $\dsection^{(1)}$ is a $sin$ function
        \item $\dsection^{(2)}$ is a $cos$ function
        \item $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$
    \end{itemize}

\begin{frame}{Sheafs}
    Restriction maps of a sheaf describe how local $\iota^*\dsection$ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}.

\subsection{Artist}
\begin{frame}{Topological Equivariant Artist Model}
The topological artist \vartist\ is a monoid equivariant sheaf map 
\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation}

where the artist $\vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)$ takes as input $\dtotal^{\prime}=\mathcal{J}^{2}(\dtotal)$. 
\end{frame}

@@

\begin{frame}{Example: Nominal Equivariance}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}
\begin{frame}{Measurement Scale Groups}

\begin{table}[H]
    \begin{tabulary}{\textwidth}{llL}
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $
    \end{tabulary}
\end{table}

    \pause
    \begin{block}{Invalid \vchannel}
     Given $\vchannel_{i}(x) = .5$ and $t(x) = x+2$,
    \pause    
    \begin{align*}
            \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + 2\\
            .5 &\neq .5 + 2
    \end{align*}
    \end{block}
\end{frame}

@@

\begin{frame}{Visualization Equivariance}
    If \vmark\ is applied to $\vsection, \vsection^{\prime}$ that generate the same \gsection\, 
    \begin{equation}
    \vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
    \end{equation}
    then the output of both sections acted on by the same monoid $m$ must be the same. 
    \pause
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}
\end{frame}

@@

\section{References}
\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}
\appendix 
\begin{frame}{Rendering: Define a Pixel}
    \begin{columns}
        \column{0.5\textwidth}
        Given a pixel
        \begin{equation}
        p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]
        \end{equation}
        the inverse map of the bounding box 
        \begin{equation}
        \gbase_{p} ={\gsection_{xy}}^{-1}(p)
        \end{equation}
        is a region $\gbase_p \subset \gbase$ such that 
        \begin{align}
            \scriptstyle r_p &= \scriptstyle \iint\limits_{S_p} \rho_r(s)ds^{2}\\
            \scriptstyle g_p &= \scriptstyle \iint\limits_{S_p} \rho_g(s)ds^{2}\\
            \scriptstyle  b_p &= \scriptstyle \iint\limits_{S_p} \rho_b(s)ds^{2}
        \end{align}
        yields the color of the pixel. 
        \column{0.5\textwidth}
        \begin{figure}[H]
            \includegraphics[width=\textwidth]{figures/math/render.png}
        \end{figure}
    \end{columns}
\end{frame}
\begin{frame}{\vfiber\ Components}

\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \label{tab:mpl_visual_variable_fiber}
\end{table}
\end{frame}
\begin{frame}[fragile]{GraphLine Data Model}
\begin{minted}{python}
class GraphLine:
    def __init__(self, FB, edge_table, vertex_table, num_samples=1000,
                        connect=False):
        #set args as attributes and generate distance
        if connect: # test connectivity if edges are continuous
            assert edge_table.keys() == self.FB.F.keys()
            assert is_continuous(vertex_table)

>>>>

\documentclass[xcolor={dvipsnames}, handout]{beamer}
%\documentclass[xcolor={dvipsnames}]{beamer}
%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}
\newtheorem{prop}{Proposition} %math?

@@

\usepackage{tabulary}
\usepackage{booktabs}
\usepackage{multirow}

\usepackage{graphicx}

@@

\usepackage[backend=bibtex, style=alphabetic, sorting=none, doi=false,isbn=false,url=false, giveninits=true]{biblatex}
\bibliography{../draft/glasslab_viz.bib}

@@

\begin{frame}{Visualizations are structure preserving maps}
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/intro/viz_same.png}
    \end{figure}
    \begin{description}
        \item[equivariance] properties of data and visual encoding match
        \item[continuity] connectivity of data and visual encoding match
        \item[composibility] structure preserved by individual components is preserved in combined components
    \end{description}
\end{frame}

\begin{frame}{Tools are tuned to continuity \cite{HeerSoftware2006}}
\begin{table}
    %\renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|L|l|l|}\toprule
        \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
        ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
        Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
        Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
         Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} & &\\
         \bottomrule
    \end{tabulary}
\end{table}
\end{frame}

\begin{frame}{Visualizations are tuned to continuinity\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007}, 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveci2012vtk}, MayaVi\cite{RamachandranMayaVI2011}, ParaView\cite{ahrens2005paraview}, Titan\cite{brianwylieUnifiedToolkitInformation2009}
    \end{enumerate}
\end{frame}

\begin{frame}{Tools are evaluated on the structure they preserve}
    \begin{description}
        \item[Naturalness] easier to understand when properties match \cite{norman_things_smart}
        \item[Graphical Integrity] graphs show \textbf{only} the data \cite{tufteVisualDisplayQuantitative2001}
        \item[Expressiveness] structure preserving mappings from data to graphic \cite{mackinlayAutomatingDesignGraphical1986}
        \item[Effectiveness] design choices made in deference to perceptual saliency \cite{clevelandResearchStatisticalGraphics1987,clevelandGraphicalPerceptionTheory1984,chambersGraphicalMethodsData1983a, munznerVisualizationAnalysisDesign2014}
    \end{description}
\end{frame}

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begin{equation}\Huge
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation}
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space.

\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation}
    \begin{description}
        \item[base space \dbase] coordinate system agnostic \textit{keys}
        \item[fiber space \dfiber] contains \textit{values}    
    \end{description}
    \begin{block}{Structural \textit{keys} with associated \textit{values} \cite{munznerVisualizationAnalysisDesign2014}}
        \begin{figure}
            \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
            \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
        \end{figure}
    \end{block}

\begin{frame}{Encode variable types in a schema like fiber \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.5\textwidth]{figures/math/fiber.png}
        \label{fig:data_fiber_example}
    \end{figure}
    \begin{block}{Fiber components can be acted on independently}
        \begin{equation}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation}
    \end{block}
\end{frame}

\begin{frame}{}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
        \caption{}
        \label{fig:base_example}
    \end{figure}
    \begin{block}{\dbase\ is an indexing space regardless of components}
        \begin{equation}
            \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
        \end{equation}
    \end{block}
\end{frame}

\begin{frame}{Data are sections\dsection\ on the \dtotal}
    For any fiber bundle, there exists a map
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
     \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
\end{frame}

\subsection{Artist}
\begin{frame}{Topological Equivariant Artist Model}
The topological artist $\vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)$ is sheaf map
\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation}
that carries homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ \cite{cegarraCohomologyMonoidsOperators2019} 
\begin{equation}\huge
\vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
\end{equation} 
\end{frame}

\begin{frame}{Existing models don't have continuity}
        \begin{description}
            \item[language model] APT, GoG: syntax, semantics, and grammar of graphics (Mackinlay, Wilkenson  \cite{mackinlayAutomatingDesignGraphical1986, mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005})
            \item[functional dependencies] constrained maps between data and visual representation(Sugibuchi \cite{sugibuchiFramwork2009}) 
            \item[category theory] data, representation, and understanding are commutative (Vickers \cite{vickersUnderstandingViz2013})
            \item[algebraic process] data ($\alpha$) and viz ($\omega$) transforms are symmetric (Kindlmann and Scheidegger \cite{kindlmannAlgebraicProcessVisualization2014})
            \begin{columns}
                \column{.5\textwidth}
                \begin{description}
                    \item[\vtotal] visual components
                    \item[$\alpha$] \monoid
                    \item[$\omega$] $\varphi(\monoid)$
                \end{description}
                \column{.5\textwidth}
                \begin{equation*}
                    \begin{tikzcd}[ampersand replacement=\&]
                        \dtotal \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& \vtotal \arrow[r, "\nu"]  \& \gtotal \arrow[d, "\omega"] \\
                        \dtotal \arrow[r, "r_2"']                     \& \vtotal \arrow[r, "\nu"'] \& \gtotal                  
                    \end{tikzcd}
                    \end{equation*}

@@

@@

\begin{frame}{Partial Order Equivariance and Stevens Scales \cite{stevensTheoryScalesMeasurement1946}}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}

\begin{table}[H]
    \begin{tabulary}{\textwidth}{llL}
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $
    \end{tabulary}
\end{table}
\end{frame}

@@

\begin{frame}{Visualization Equivariance}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}
    \begin{prop}
        If for elements of the monoid $m \in \monoid$ and for all $\vsection, \vsection^{\prime} \in X$, we define the monoid action on $X$ so that it is by definition equivariant
        \begin{equation}
        \vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
        \label{eq:math:artist:q}
        \end{equation}
        then a monoid action on $Y$ can be defined as $m\circ \gsection = \gsection^{\prime}$. If and only if \vmark\ satistfies \autoref{eq:math:artist:q}, we can state that the transformed graphic $\gsection^{\prime}=\vmark(m\circ \vsection)$ is equivariant to a monoid action applied on \vmark\ with input $\vsection \in \vmark^{-1}(\gsection)$ that must generate valid \gsection. 
    \end{prop}
\end{frame}

@@

\section{References}
\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}
\appendix 
\begin{frame}{Fiber is all possible values a variable can be \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
    Given a space of all possible values \ftotal\
    \begin{equation}
        \label{eq:data_types}
        \begin{tikzcd}[ampersand replacement=\&]
            \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] \& \ftotal \arrow[d, "\pi"] \\
            \fnames \arrow[r, "\fsection"']                  \& \ftypes       
        \end{tikzcd}
    \end{equation}
    a fiber component is the restricted space $\ftotal_{\fsection(\fname)}$. 
    \begin{equation}
        \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
    \end{equation}
    \begin{description}
        \item[\ftypes] data types of the variables in the dataset 
        \item[\ftotal] disjoint union of all values of type $\ftype \in \ftypes$ 
        \item[\fnames] variable names, $\fname \in \fnames$
        \item[\fttype] \ftotal\ restricted to the data type of a named variable   
    \end{description}
\end{frame}
\begin{frame}{Monoid actions}
    A monoid \monoid\ is a set with
    \begin{description}
        \item[associative binary operator] $\ast:\monoid \times \monoid\rightarrow \monoid$
        \item[identity element] $e\in \monoid$ such that $e\ast a= a \ast e = a$ for all $a \in \monoid$. 
    \end{description}
    \pause
    \begin{block}{left monoid action}
    A set \dfiber\ with an action\cite{nlab:action} $\bullet: \monoid\times \dfiber \rightarrow \dfiber$ with the properties:
        \begin{align*}
            \textbf{associativity}\;& \text{for all } f,g \in \monoid \text{ and } x\in \dfiber,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
            \textbf{identity}\;& \text{for all } x\in \dfiber, e\in \monoid,\,  e\bullet x = x 
        \end{align*}
    \end{block}
\end{frame}
\begin{frame}{Keeping track of sections with sheafs}
    Restriction maps of a sheaf describe how local $\iota^*\dsection$ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}.

\begin{frame}{Rendering: Define a Pixel}
    \begin{columns}
        \column{0.5\textwidth}
        Given a pixel
        \begin{equation}
        p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]
        \end{equation}
        the inverse map of the bounding box 
        \begin{equation}
        \gbase_{p} ={\gsection_{xy}}^{-1}(p)
        \end{equation}
        is a region $\gbase_p \subset \gbase$ such that 
        \begin{align}
            \scriptstyle r_p &= \scriptstyle \iint\limits_{S_p} \rho_r(s)ds^{2}\\
            \scriptstyle g_p &= \scriptstyle \iint\limits_{S_p} \rho_g(s)ds^{2}\\
            \scriptstyle  b_p &= \scriptstyle \iint\limits_{S_p} \rho_b(s)ds^{2}
        \end{align}
        yields the color of the pixel. 
        \column{0.5\textwidth}
        \begin{figure}[H]
            \includegraphics[width=\textwidth]{figures/math/render.png}
        \end{figure}
    \end{columns}
\end{frame}
\begin{frame}{\vfiber\ Components}

\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \label{tab:mpl_visual_variable_fiber}
\end{table}
\end{frame}
